### 2017-06-12 paper.tex
reversed frequency preference revised

A classic experiment by DeKonick and Schulman~\cite{DeKoninck:1998wh} showed that \textit{in vitro} CaMKII holoenzymes activation are sensitive to the frequency of Ca\textsuperscript{2+}-CaM pulses. A recent study \textit{in vivo} also noticed glutamate uncaging frequency affects CaMKII activation in spine~\cite{Fujii:2013bg}. Numerous modeling studies have also reported a dependence of CaMKII activation on the frequency of calcium signals~\cite{}. To confirm this CaMKII activation pattern, we tested our 6-subunit CaMKII model using the previously generated 10 Hz and 5 Hz Ca\textsuperscript{2+} influx files (see Methods). The total number of influx Ca\textsuperscript{2+} ions are comparable (40433 ions in 5 Hz vs 40435 ions in 10 Hz). For each frequency, we simulated 10 trials and presented the time-varying mean +/- standard deviation of phosphorylated subunits. As shown in Fig~\ref{fig}, we observed significantly higher phosphorylation levels with 10 Hz than with 5 Hz input.

The observed frequency preference for 10 Hz is inherent to the Ca\textsuperscript{2+}-CaM-CaMKII binding. Through examining the reaction occurrences over time, we noticed that the 10 Hz input results in more bindings between NxC2 CaM and CaMKII, and consequently more autophosphorylations (Fig~\ref{}). Furthermore, the frequency effect can be modulated by slowing down Ca\textsuperscript{2+} diffusion and limiting the total amount of CaM. In the presence of moderate amount of Ca\textsuperscript{2+}, CaM molecules follow the predominant pathway by choosing reaction paths according to affinities. As a result, CaMKII interactions with CaM and Ca\textsuperscript{2+} binding to CaM are relatively independent. In contrast, when Ca\textsuperscript{2+} influx becomes saturating, the flush of Ca\textsuperscript{2+} quickly takes up the CaM, leaving CaMKII temporarily lack of binding partners. This temporary shortage of CaM results in an alternative pathway decision and ultimately a suboptimal phosphorylation level. Given the same total amount of Ca\textsuperscript{2+}, 10 Hz is easier to saturate CaM in a short time than 5 Hz. To demonstrate the effect of saturating Ca\textsuperscript{2+}, we increased the Ca\textsuperscript{2+} channel number from 16 channels as in the prototype model to allow more Ca\textsuperscript{2+} ions to flow in per action potential (Fig~\ref{fig5}). In the mean time, the total Ca\textsuperscript{2+} influx is kept consistent for 5 Hz and 10 Hz. As Ca\textsuperscript{2+} level increases, the network produces more phosphorylated CaMKII subunits, and the phosphorylation level difference between 5 Hz and 10 Hz conditions diminishes. Eventually, the 10 Hz input becomes saturating and the network generates less phosphorylation under 10 Hz than the 5 Hz stimulus (Fig~\ref{fig5}F).

@@

We examined the reaction occurrences profile over time on Layer1 edges (Fig~\ref{fig6}) and noticed that the reaction pathway choice deviates from the predominant pathway when Ca\textsuperscript{2+} level becomes saturating. This trend can be revealed from the pathway choice for CaM N0C2 (Fig~\ref{fig6}A-C). At nonsaturating Ca\textsuperscript{2+} level, N0C2 molecules are more likely to bind with CaMKII (reaction KN0C2) than binding a Ca\textsuperscript{2+} ion on the N1 site (reaction N1). When Ca\textsuperscript{2+} becomes saturating, N0C2 molecules tends to choose N1 binding over reaction KN0C2. This switch of reaction choice is observed for 10 Hz but not for 5 Hz. Likewise, we analyzed the profile of all possible CaMKII-CaM binding reactions. For 5 Hz conditions, binding of N0C2 consistently dominates over other CaMKII-CaM binding reactions, regardless of the amount of Ca\textsuperscript{2+} influx (Fig~\ref{fig6}D-F). However for 10 Hz, as Ca\textsuperscript{2+} influx increases, binding of N2C2 (reaction KN2C2) gradually replaces binding of N0C2 (reaction KN0C2) and becomes the dominant choice for CaM to enter Layer2 (Fig~\ref{fig6}G-I). In breif, this switch of dominant pathway occurs for 10 Hz, correlating with the suboptimal phosphorylation level observed for 10 Hz.

@@

The saturating Ca\textsuperscript{2+} effects on frequency preference can also be facilitated by slowing Ca\textsuperscript{2+} diffusion and limiting the available CaM in the system (Fig~\ref{fig7}). A slowed Ca\textsuperscript{2+} diffusion allows more interactions between Ca\textsuperscript{2+} and CaM, effectively mimicking a saturating level Ca\textsuperscript{2+}. To demonstrate, we set Ca\textsuperscript{2+} diffusion as \SI{0.0000011}{\square\cm\per\s}, which is half of the default speed and does not change the activation-limited regime of the network. Slow diffusion results in a dramatic increase of CaMKII phosphorylation regardless of Ca\textsuperscript{2+} channels and input frequencies. Also, the frequency preference for 10 Hz is reversed at $3\times$ of $6\times$ Ca\textsuperscript{2+} influx condition(Fig~\ref{fig7}C,D). In other words, with slow diffusion, Ca\textsuperscript{2+} influx provided by $3\times$ becomes sufficiently saturating to change the frequency preference. On the other hand, the same effect can be artificially created by limiting the total amount of CaM in the system (Fig~\ref{fig7}A,B). To demonstrate, we compared simulations using $3.5\times$ Ca\textsuperscript{2+} input with the default amount of CaM (\SI{5}{\micro\Molar}) and limited CaM (\SI{2.5}{\micro\Molar}). Even though the Ca\textsuperscript{2+} amount does not change, by limiting CaM we noticed a reverse of frequency preference from 10 Hz to 5 Hz. It is also worth noting is that by combining slow diffusion and limited CaM, reversing the frequency preference is possible with merely twice amount of Ca\textsuperscript{2+} influx (Fig~\ref{fig7}E,F).

>>>>

A classic experiment by DeKonick and Schulman~\cite{DeKoninck:1998wh} showed that \textit{in vitro} CaMKII holoenzymes activation are sensitive to the frequency of Ca\textsuperscript{2+}-CaM pulses. A recent study also showed that \textit{in vivo} glutamate uncaging frequency affects CaMKII activation in spine~\cite{Fujii:2013bg}. Numerous modeling studies have also reported a dependence of CaMKII activation on the frequency of calcium signals~\cite{}. To confirm this CaMKII activation pattern, we tested our 6-subunit CaMKII model using the previously generated 10 Hz and 5 Hz Ca\textsuperscript{2+} influx files (see Methods). The total number of influx Ca\textsuperscript{2+} ions are comparable (40433 ions in 5 Hz vs 40435 ions in 10 Hz). For each frequency, we simulated 10 trials and presented the time-varying mean +/- standard deviation of phosphorylated subunits. As shown in Fig~\ref{fig}, we observed significantly higher phosphorylation levels with 10 Hz than with 5 Hz input.

We examined the reaction occurrences over time and noticed that the 10 Hz input results in more bindings between NxC2 CaM and CaMKII, and consequently more autophosphorylations (Fig~\ref{}). Therefore the frequency effect is inherent to the binding between Ca\textsuperscript{2+}, CaM and CaMKII. The frequency effect can be reversed by providing saturating amount of Ca\textsuperscript{2+}. To demonstrate, we increased the Ca\textsuperscript{2+} channel number to allow more Ca\textsuperscript{2+} influx per action potential (Fig~\ref{fig5}). In the mean time, the total Ca\textsuperscript{2+} influx is kept consistent for 5 Hz and 10 Hz. As Ca\textsuperscript{2+} level increases, the network produces more phosphorylated CaMKII subunits for both input frequencies, but the phosphorylation level difference between 5 Hz and 10 Hz diminishes. Eventually, the 10 Hz input becomes saturating and the network generates less phosphorylation under 10 Hz than the 5 Hz input (Fig~\ref{fig5}F).

Equivalently, the saturating Ca\textsuperscript{2+} effects on frequency preference can also be facilitated by limiting the available CaM and slowing Ca\textsuperscript{2+} diffusion (Fig~\ref{fig7}). To demonstrate, we compared simulations using $3.5\times$ Ca\textsuperscript{2+} input with the default amount of CaM (\SI{5}{\micro\Molar}) and limited CaM (\SI{2.5}{\micro\Molar}). Even though the Ca\textsuperscript{2+} input does not change, by limiting CaM we noticed a reverse of frequency preference from 10 Hz to 5 Hz. The same effect can be achieved with slowed Ca\textsuperscript{2+} diffusion. We set Ca\textsuperscript{2+} diffusion as \SI{0.0000011}{\square\cm\per\s}, which is half of the default speed and does not change the activation-limited regime of the network. Slow diffusion results in a dramatic increase of CaMKII phosphorylation regardless of Ca\textsuperscript{2+} amount and input frequencies. Also, the frequency preference for 10 Hz is reversed at $3\times$ Ca\textsuperscript{2+} influx condition(Fig~\ref{fig7}C,D). In other words, with slow diffusion, Ca\textsuperscript{2+} influx provided by $3\times$ becomes sufficiently saturating to change the frequency preference. It is also worth noting is that by combining limited CaM and slow diffusion, reversing the frequency preference is possible with merely twice amount of Ca\textsuperscript{2+} influx (Fig~\ref{fig7}E,F).

To explain what happened, we profiled the reaction occurrences over time on Layer1 edges. We noticed that the reaction pathway choice deviates from the predominant pathway when Ca\textsuperscript{2+} level becomes saturating. In the presence of moderate amount of Ca\textsuperscript{2+}, CaM molecules follow the predominant pathway by choosing reaction paths according to affinities. On one hand, bindings between N0C2 and CaMKII (rxn_KN0C2) are more likely to occur than bindings of Ca\textsuperscript{2+} ions on N1 sites of CaM molecules (rxn_N1). Similarly, CaMKII binding with N0C2 (rxn_KN0C2) dominates all types of CaMKII-CaM binding reactions (Fig~\ref{fig}). However, when \textsuperscript{2+} influx becomes saturating, CaM molecules tend to stay in the Layer1 till they get fully loaded by Ca\textsuperscript{2+}. As a result, CaMKII binding with N2C2 (rxn_KN2C2) becomes dominant. This decision change gives rise to an altered pathway and eventually a suboptimal phosphorylation level. Given the same total amount of Ca\textsuperscript{2+}, 10 Hz is easier to saturate CaM in a short time than 5 Hz. Correspondingly, the switch of pathway choice is more dramatic for 10 Hz at $4.5\times$ Ca\textsuperscript{2+}, resulting in the revserved frquency perference as observed previously.

@@

\begin{comment}
We examined the reaction occurrences profile over time on Layer1 edges (Fig~\ref{fig6}) and noticed that the reaction pathway choice deviates from the predominant pathway when Ca\textsuperscript{2+} level becomes saturating. This trend can be revealed from the pathway choice for CaM N0C2 (Fig~\ref{fig6}A-C). At nonsaturating Ca\textsuperscript{2+} level, N0C2 molecules are more likely to bind with CaMKII (reaction KN0C2) than binding a Ca\textsuperscript{2+} ion on the N1 site (reaction N1). When Ca\textsuperscript{2+} becomes saturating, N0C2 molecules tends to choose N1 binding over reaction KN0C2. This switch of reaction choice is observed for 10 Hz but not for 5 Hz. Likewise, we analyzed the profile of all possible CaMKII-CaM binding reactions. For 5 Hz conditions, binding of N0C2 consistently dominates over other CaMKII-CaM binding reactions, regardless of the amount of Ca\textsuperscript{2+} influx (Fig~\ref{fig6}D-F). However for 10 Hz, as Ca\textsuperscript{2+} influx increases, binding of N2C2 (reaction KN2C2) gradually replaces binding of N0C2 (reaction KN0C2) and becomes the dominant choice for CaM to enter Layer2 (Fig~\ref{fig6}G-I). In breif, this switch of dominant pathway occurs for 10 Hz, correlating with the suboptimal phosphorylation level observed for 10 Hz.
\end{comment}

@@
